A Cross-Domain Benchmark for Active Learning

### 2024-06-05 doc/al_benchmark_paper.tex
paper version 1.0

Active Learning (AL) deals with identifying the most informative samples for
labeling to reduce data annotation costs for supervised learning tasks. AL
research suffers from the fact that lifts from literature generalize poorly and
that only a small number of repetitions of experiments are conducted. To overcome
these obstacles, we propose \emph{CDALBench}, the first active learning benchmark
which includes tasks in computer vision, natural language processing and tabular
learning. Furthermore, by providing an efficient, greedy oracle, \emph{CDALBench}
can be evaluated with 50 runs for each experiment. We show, that both the
cross-domain character and a large amount of repetitions are crucial for
sophisticated evaluation of AL research. Concretely, we show that the
superiority of specific methods varies over the different domains, making it
important to evaluate Active Learning with a cross-domain benchmark.
Additionally, we show that having a large amount of runs is crucial. With only
conducting five runs as often done in the literature, the superiority of
specific methods can strongly vary with the specific runs. This effect is so strong, that, depending on the seed, even a well-established method's performance can be significantly better and significantly
worse than random for the same dataset. 
\end{abstract}

@@

Here we step in with \emph{CDALBench}, an AL benchmark which covers
multiple application domains and reports a large amount of runs per
experiment, so that the significance of performance differences can can be
estimated. Specifically, \emph{CDALBench} consists of datasets from
computer vision, natural language processing and the tabular domain. We provide
our datasets both in normal format %(i.e. consisting of text, images etc)
as well as ``embedded'' by a fixed embedding model, enabling evaluation of AL methods in
this semi-supervised setting.
Furthermore, we propose two novel synthetic datasets to highlight general
challenges for AL methods.
\begin{table}
	\centering
	\caption{Comparison of our benchmark with the existing literature. Oracle curves
		serve as an approximation of the best possible AL algorithm. Our benchmark contains 9 
		datasets (14 including the encoded versions). ``Semi"
		indicates whether the paper is employing any form of self- or semi-supervised
		learning. A ``-" for repetitions means that we could not determine how often each
		experiment was repeated in the respective framework. \emph{CDALBench} is the only
		benchmark which evaluates a high number of runs and considers
		all 5 domains.}
	\label{tab:benchmark_comparison}
	\resizebox{\columnwidth}{!}{
		\begin{tabular}{lccc|cllll|cl}
			Paper & Sampling & \#Data & \#Alg & \multicolumn{1}{l}{Img} & Txt & Tab & Synth & Semi & Oracle & Repetitions \\ \hline
			\multicolumn{1}{l|}{Beck et al. \cite{beck2021effective}} & batch & 4 & 7 & \checkmark & - & - & - & - & - & - \\
			\multicolumn{1}{l|}{Hu et al. \cite{hu2021towards}} & batch & 5 & 13 & \checkmark & \checkmark & - & - & - & - & 3 \\
			\multicolumn{1}{l|}{Zhou et al. \cite{zhou2021towards}} & batch & 3 & 2 & \checkmark & \checkmark & - & - & - & \checkmark & 5 \\
			\multicolumn{1}{l|}{Zhan et al. \cite{zhan2021comparative}} & sngl+batch & 35 & 18 & - & - & \checkmark & \checkmark & - & \checkmark & 10-100 \\
			\multicolumn{1}{l|}{Munjal et al. \cite{munjal2022towards}} & batch & 2 & 8 & \checkmark & - & - & - & - & - & 3 \\
			\multicolumn{1}{l|}{Li et al. \cite{li2022empirical}} & batch & 5 & 13 & \checkmark & - & - & - & \checkmark & - & - \\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
While multiple benchmark suites have been proposed for AL, none of them provide experiments for more than two domains.
The authors of \cite{beck2021effective}, \cite{munjal2022towards}, \cite{li2022empirical}, \cite{ji2023randomness} and \cite{luth2024navigating} even focus exclusively on the image domain.
Especially the tabular domain is underrepresented in preceding benchmarks, as only \cite{zhan2022comparative} provides experiments for it.
Experiments on the interplay between AL and semi-supervised learning are similarly under-researched, as only two works exist \cite{li2022empirical, luth2024navigating}, both of them only using images.
An oracle algorithm has been proposed by two works \cite{zhou2021towards, zhan2022comparative}. 
Both of these algorithms rely on search and are computationally very expensive, while our proposed method efficiently can be constructed sequentially.
%While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call ``Label Efficiency'' and provide experiments on many common configurations of data preparation, model training, and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning.
%The authors of \cite{hu2021towards} study models that are trained on actively learned datasets in the image and text domain.
%They test for several different properties of the models including robustness, response to compression techniques and final performance.
%\cite{zhou2021towards} proposed an oracle algorithm for AL that uses Simulated Annealing search to approximate a solution for the optimal subset of labeled data.
%Additionally, they study the generalization behavior of subsets of labeled data in the text an image domain.
The two closest related works to this benchmark are \cite{ji2023randomness} and \cite{luth2024navigating}, who also place a much higher emphasis on the problem of evaluating AL methods under high variance than their predecessors (indicated in Tab. \ref{tab:benchmark_comparison} by a dashed line).
The authors of \cite{ji2023randomness} posed a total of 11 ``recommendations" for reliable evaluation of AL methods.
We largely adapt the proposed recommendations and extend their work to multiple domains and query sizes.
For a complete list of the recommendations and our implementation of them, please refer to App. \ref{app:recommendations}.
This work also pays attention to the so-called ``pitfalls" of AL evaluation proposed in \cite{luth2024navigating}.
For a complete list of the pitfalls and our considerations regarding them, please refer to App. \ref{app:pitfalls}.
To the best of our knowledge, we are the first to extend reliable SOTA (based on \cite{ji2023randomness, luth2024navigating}) experimentation to a total of 5 data domains and a high number of repetitions per experiment.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Seeding vs. Repetitions}\label{sec:reproducibility}
Considering the high computational cost of 50 repetitions, another approach to ensure consistency between experiments would be to reduce the amount of variance in the experiment by keeping as many subsystems (weight initialization, data splits, etc.) as possible fixed with specialized seeding. \\
We describe a novel seeding strategy in Appendix \ref{app:seeding_strategy} that is capable of tightly controlling the amount variance in the experiment.
However, previous works have noted that an actively sampled, labeled set does not generalize well between model architectures or even different initializations of the same model (\cite{zhou2021towards, lowell2018practical}), providing a bad approximation of the quality of an AL method (i.e. measured performances for an method might not even transfer to a different model initialization).
Hence, we opt for letting the subsystems vary in controlled way (For details, please refer to App. \ref{app:seeding_strategy}) and combine that with a high number of repetitions to obtain a good average of the generalization performance of each AL method.

@@

			Enc. DNA    & Linear & 40 & o & o &&&&& \\
			Enc. Splice & Linear & 100 & o & o & o & o &&&\\

@@

			Enc. Cifar10& Linear & 450 & o & o & o & o & o && \\
			Enc. FMnist & Linear & 500 & o & o & o & o & o && \\
			Enc. USPS   & Linear & 600 & o & o & o & o & o && \\

@@

However, we are solely focused on comparing different AL methods in this paper and do not aim to develop novel classification models on these datasets.
Our assumption is that a well-performing method in our benchmark will also generalize well to larger real-world datasets, because we included multiple different data domains, classifier types and sizes in our experiments. \\ [1mm]
Adapting the semi-supervised setting from \cite{hacohen2022active}, we offer all our datasets un-encoded (normal) as well as pre-encoded by a fixed embedding model that was trained by unsupervised contrastive learning (semi-supervised).

@@

Every dataset has a fixed size for the seed set of 1 sample per class, with the only exceptions being un-encoded FashionMnist and Cifar10 with 100 examples per class to alleviate the cold-start problem in these complex domains.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation Protocol}\label{sec:evaluation}
Following \cite{zhou2021towards}, the quality of an AL method is evaluated by an ``anytime protocol" that incorporates classification performance at every iteration, as opposed to evaluating final performance after the budget is exhausted.
We employ the normalized area under the accuracy curve (AUC):
\begin{equation}\label{eq:auc}
	\operatorname{AUC}(\D_\test, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(\D_\test, \hat y^{(i)})
\end{equation}
%The AUC incorporates performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
%AL algorithms have to perform well in both scenarios. \\ [1mm]
Since AUC is still influenced by the budget, we define a set of rules to set this hyperparameter upfront, so that we are not favoring a subset of methods by handcrafting a budget.
In this work, we choose the budget per dataset to be the first point at which one of 2 stopping conditions apply: (i) an method (except oracle) manages to reach 99\% of the full-dataset-performance (using the smallest query size) or (ii) the best method (except oracle) did not improve the classifier's accuracy by at least 2\% in the last 20\% of iterations.
The first rule follows \cite{ji2023randomness}, while the second rule prevents excessive budgets for cases with diminishing returns in the budget.
The resulting budgets can be found in Tab. \ref{tab:batch_sizes}. \\ [1mm]
As described in Sec. \ref{sec:restarts}, we repeat each experiment multiple times.
Each repetition retains the train/test split (often given by the dataset itself), but creates a new validation split that is sampled (based on $s_\D$) from the entire dataset (not just the seed set $\LL^{(0)}$). \\ [1mm]
%
Apart from plotting standard performance curves and reporting their AUC values per dataset in App. \ref{app:all_results}, we primarily rely on ranks to aggregate the performance of an AL method across datasets.
For each dataset and query size, the AUC values of all AL methods are sorted and assigned a rank based on position, with the best rank being 1.
These ranks can safely be averages across datasets as they are no longer subjected to scaling differences of each dataset.
Additionally, we employ Critical Difference (CD) diagrams (like Fig. \ref{fig:ranks_by_domain}) for statistical testing.
CD diagrams \cite{IsmailFawaz2018deep} use the Wilcoxon signed-rank test, which is a variant of the paired T-test, to find significant differences of ranks between AL methods.
For a detailed description of how every CD diagram is created, please refer to App. \ref{app:cd_diagrams}.

@@

Using additional resources, like excessive training time, or direct access to the labeled test set, an oracle method for AL finds the oracle set $\mathcal{O}_b$ for a given dataset, model, and

@@

However, due to the combinatorial nature of the problem, it is computationally

@@

annealing to search for a subset with maximal test accuracy and used the best solution found

@@

uniformly samples at iteration $i$ a subset $\mathcal{U}_S$ of the not already

@@

point $u$ for which the classifier $\hat{y}$ trained on

@@

our oracle frequently encounters situations where every point in $u_k$ would

@@

defaults to margin sampling in that iteration. The resulting greedy algorithm

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Implementation Details}\label{sec:implementation_details}
At each iteration $i$ the AL method $\Omega$ picks an unlabeled datapoint based on a fixed set of information $\{\mathcal{L}^{(i)}, \mathcal{U}^{(i)}, B, |\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|, \text{acc}^{(i)}, \text{acc}^{(1)}, \hat y^{(i)}, \text{opt}_{\hat y}\}$, where $\text{opt}_{\hat y}$ is the optimizer used to fit $\hat y^{(i)}$.
This set grants full access to the labeled and unlabeled set, as well as all parameters of the classifier and the optimizer.
Additionally, we provide meta-information, like the size of the seed set through $|\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|$, the remaining budget though the addition of $B$ and the classifiers potential through $\text{acc}^{(1)}$ and $\text{acc}^{(i)}$.
We allow AL methods to derive information from this set, e.g. predictions of the classifier $\hat y^{(i)}(x); \hspace{2mm} x \in \mathcal{U}^{(i)} \cup \mathcal{L}^{(i)}$, clustering, or even training additional models.
However, the method may not incorporate external information e.g. other datasets, queries to recover additional labels, additional training steps for $\hat y$, or the test/validation set. \\
For our study we selected acquisition functions with good performances reported by multiple different sources that can work with the set of information stated above.
For a list of all AL methods, please refer to Table \ref{tab:results}, with detailed descriptions being found in Appendix \ref{app:acquisition_functions}. \\ [1mm]
%
The model $\hat y$ can be trained in two ways. Either the parameters of the model are reset to a fixed initial setting $\hat y^{(0)}$ after each AL iteration and the classifier is trained from scratch with the updated labeled set $\mathcal{L}^{(i)}$, or the previous state $\hat y^{(i-1)}$ is retained and the classifier is fine-tuned on $\mathcal{L}^{(i)}$ for a reduced number of epochs.
In this work, we use the fine-tuning method for un-encoded datasets to save computational time, while we use the from-scratch training for embedded datasets since they have very small classifiers and this approach generally produces better results.
Our fine-tuning scheme always trains for at least one epoch and employs an aggressive early stopping with a patience of 2 afterwards.
%This work employs a validation split that is produced from the source dataset itself before active learning instead of splitting it from the seed set or omitting it altogether.
%Even though the use of a fully labeled validation set might be regarded as impractical, since such a set will never exist during deployment, we strongly advocate for using it in \textbf{research}, as it can be used for hyperparameter tuning, resulting in more stable hyperparameters and reducing the overall training stochasticity.
%We optimize all hyperparameters on our fully labeled validation sets using grid search.

@@

Please note, that we are averaging not only over runs, but also over query sizes per dataset. 
This leads to a negative impact for AL methods that do not adapt well to larger query sizes.

@@

As mentioned in contribution \ref{cont:domains}, our results on real-world data show significant differences in the performance of methods between data domains.

@@

Results for the semi-supervised domain appear mostly in line with the other 3 domains, but a closer analysis of performances split into semi-sup. images and semi-sup. tabular reveals the need for further research.

@@

Both types of methods have principled shortcomings in terms of the utilized information that makes them unsuitable for certain data distributions. 

@@

To avoid methods memorizing these datasets they are generated from scratch for each experiment, depending on $s_\D$. \\

@@

The honeypot is located on the likely decision boundary of a classifier that is trained on the beneficial samples to maximize its negative impact on purely uncertainty-based AL methods.
Diverging Sine samples the datapoints for each class from two diverging sinusoidal functions that are originating from the same y-intercept.

@@

Both margin sampling and BADGE (the two best methods) being vulnerable to adverse samples or simply measurement noise, highlights the need for further research in this area. \\

>>>>

Active Learning (AL) deals with identifying the most informative samples for
labeling to reduce data annotation costs for supervised learning tasks. AL
research suffers from the fact that lifts from literature generalize poorly and
that only a small number of repetitions of experiments are conducted. To overcome
these obstacles, we propose \emph{CDALBench}, the first active learning benchmark
which includes tasks in computer vision, natural language processing and tabular
learning. Furthermore, by providing an efficient, greedy oracle, \emph{CDALBench}
can be evaluated with 50 runs for each experiment. We show, that both the
cross-domain character and a large amount of repetitions are crucial for
sophisticated evaluation of AL research. Concretely, we show that the
superiority of specific methods varies over the different domains, making it
important to evaluate Active Learning with a cross-domain benchmark.
Additionally, we show that having a large amount of runs is crucial. With only
conducting three runs as often done in the literature, the superiority of
specific methods can strongly vary with the specific runs. This effect is so strong, that, depending on the seed, even a well-established method's performance can be significantly better and significantly
worse than random for the same dataset. 
\end{abstract}

@@

Here we step in with \emph{CDALBench}, an AL benchmark which covers
multiple application domains and reports a large amount of runs per
experiment, so that the significance of performance differences can be
estimated. Specifically, \emph{CDALBench} consists of datasets from
computer vision, natural language processing and the tabular domain. We provide
our datasets both in normal format %(i.e. consisting of text, images etc)
as well as ``embedded'' by a fixed embedding model, enabling evaluation of AL methods in
this semi-supervised setting.
Furthermore, we propose two novel synthetic datasets to highlight general
challenges for AL methods.
\begin{table}
	\centering
	\caption{Comparison of our benchmark with the existing literature. Oracle curves
		serve as an approximation of the best possible AL algorithm. Our benchmark contains 9 
		datasets (14 including the encoded versions). ``Semi"
		indicates whether the paper is employing any form of self- or semi-supervised
		learning. A ``-" for repetitions means that we could not determine how often each
		experiment was repeated in the respective framework. \emph{CDALBench} is the only
		benchmark which evaluates a high number of runs and considers
		all 5 domains.}
	\label{tab:benchmark_comparison}
	\resizebox{\columnwidth}{!}{
		\begin{tabular}{lccc|cllll|cl}
			Paper & Sampling & \#Data & \#Alg & \multicolumn{1}{l}{Img} & Txt & Tab & Synth & Semi & Oracle & Repetitions \\ \hline
			\multicolumn{1}{l|}{Beck et al. \cite{beck2021effective}} & batch & 4 & 7 & \checkmark & - & - & - & - & - & - \\
			\multicolumn{1}{l|}{Hu et al. \cite{hu2021towards}} & batch & 5 & 13 & \checkmark & \checkmark & - & - & - & - & 3 \\
			\multicolumn{1}{l|}{Zhou et al. \cite{zhou2021towards}} & batch & 3 & 2 & \checkmark & \checkmark & - & - & - & \checkmark & 5 \\
			\multicolumn{1}{l|}{Zhan et al. \cite{zhan2021comparative}} & sngl+batch & 35 & 18 & - & - & \checkmark & \checkmark & - & \checkmark & 10-100 \\
			\multicolumn{1}{l|}{Munjal et al. \cite{munjal2022towards}} & batch & 2 & 8 & \checkmark & - & - & - & - & - & 3 \\
			\multicolumn{1}{l|}{Li et al. \cite{li2022empirical}} & batch & 5 & 13 & \checkmark & - & - & - & \checkmark & - & - \\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
While multiple benchmark suites have been proposed for AL, none of them provide experiments for more than two domains.
The authors of \cite{beck2021effective}, \cite{munjal2022towards}, \cite{li2022empirical}, \cite{ji2023randomness} and \cite{luth2024navigating} even focus exclusively on the image domain.
Especially the tabular domain is underrepresented in preceding benchmarks, as only \cite{zhan2022comparative} provides experiments for it.
The interplay between AL and semi-supervised learning is similarly under-researched, as only two works exist \cite{li2022empirical, luth2024navigating}, both of them only using images.
An oracle algorithm has been proposed by two works \cite{zhou2021towards, zhan2022comparative}. 
Both of these algorithms rely on search and are computationally very expensive, while our proposed method efficiently can be constructed sequentially.
%While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call ``Label Efficiency'' and provide experiments on many common configurations of data preparation, model training, and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning.
%The authors of \cite{hu2021towards} study models that are trained on actively learned datasets in the image and text domain.
%They test for several different properties of the models including robustness, response to compression techniques and final performance.
%\cite{zhou2021towards} proposed an oracle algorithm for AL that uses Simulated Annealing search to approximate a solution for the optimal subset of labeled data.
%Additionally, they study the generalization behavior of subsets of labeled data in the text an image domain.
The two closest related works to this benchmark are \cite{ji2023randomness} and \cite{luth2024navigating}, who also place a much higher emphasis on the problem of evaluating AL methods under high variance than their predecessors (indicated in Tab. \ref{tab:benchmark_comparison} by a dashed line).
The authors of \cite{ji2023randomness} posed a total of 11 ``recommendations" for reliable evaluation of AL methods.
We largely adapt the proposed recommendations and extend their work to multiple domains and query sizes.
For a complete list of the recommendations and our implementation of them, please refer to App. \ref{app:recommendations}.
This work also pays attention to the so-called ``pitfalls" of AL evaluation proposed in \cite{luth2024navigating}.
For a complete list of the pitfalls and our considerations regarding them, please refer to App. \ref{app:pitfalls}.
To the best of our knowledge, we are the first to extend reliable SOTA (based on \cite{ji2023randomness, luth2024navigating}) experimentation to a total of 5 data domains and a high number of repetitions per experiment.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Seeding vs. Repetitions}\label{sec:reproducibility}
Considering the high computational cost of 50 repetitions, another approach to ensure consistency between experiments would be to reduce the amount of variance in the experiment by keeping as many subsystems (weight initialization, data splits, etc.) as possible fixed with specialized seeding. \\
We describe a novel seeding strategy in Appendix \ref{app:seeding_strategy} that is capable of tightly controlling the amount variance in the experiment.
However, previous works have noted that an actively sampled, labeled set does not generalize well between model architectures or even different initializations of the same model (\cite{zhou2021towards, lowell2018practical}), providing a bad approximation of the quality of an AL method (i.e. measured performances for an AL method might not even transfer to a different model initialization).
Hence, we opt for letting the subsystems vary in controlled way (For details, please refer to App. \ref{app:seeding_strategy}) and combine that with a high number of repetitions to obtain a good average of the generalization performance of each AL method.

@@

			Semi DNA    & Linear & 40 & o & o &&&&& \\
			Semi Splice & Linear & 100 & o & o & o & o &&&\\

@@

			Semi Cifar10& Linear & 450 & o & o & o & o & o && \\
			Semi FMnist & Linear & 500 & o & o & o & o & o && \\
			Semi USPS   & Linear & 600 & o & o & o & o & o && \\

@@

We are solely focused on comparing different AL methods in this paper and do not aim to develop novel classification models on these datasets.
Our assumption is that a well-performing method in our benchmark will also generalize well to larger datasets and classifiers, because we included multiple different data domains, classifier types and sizes in our experiments. \\ [1mm]
Adapting the semi-supervised setting from \cite{hacohen2022active}, we offer all our datasets un-encoded (normal) as well as pre-encoded (semi-supervised) by a fixed embedding model that was trained by unsupervised contrastive learning .

@@

Every dataset has a fixed size for the seed set $\LL^{(0)}$ of 1 sample per class, with the only exceptions being un-encoded FashionMnist and Cifar10 with 100 examples per class to alleviate the cold-start problem in these complex domains.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation Protocol}\label{sec:evaluation}
Following \cite{zhou2021towards}, the quality of an AL method is evaluated by an ``anytime protocol" that incorporates classification performance at every iteration, as opposed to evaluating final performance after the budget is exhausted.
We employ the normalized area under the accuracy curve (AUC):
\begin{equation}\label{eq:auc}
	\operatorname{AUC}(\D_\test, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(\D_\test, \hat y^{(i)})
\end{equation}
%The AUC incorporates performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
%AL algorithms have to perform well in both scenarios. \\ [1mm]
Since AUC is still influenced by the budget, we define a set of rules to set this hyperparameter upfront, so that we are not favoring a subset of methods by handcrafting a budget.
In this work, we choose the budget per dataset to be the first point at which one of 2 stopping conditions apply: (i) an method (except oracle) manages to reach 99\% of the full-dataset-performance (using the smallest query size) or (ii) the best method (except oracle) did not improve the classifier's accuracy by at least 2\% in the last 20\% of iterations.
The first rule follows \cite{ji2023randomness}, while the second rule prevents excessive budgets for cases with diminishing returns in the budget.
The resulting budgets can be found in Tab. \ref{tab:batch_sizes}. \\ [1mm]
As described in Sec. \ref{sec:restarts}, we repeat each experiment 50 times.
Each repetition retains the train/test split (often given by the dataset itself), but creates a new validation split that is sampled from the entire dataset (not just the seed set $\LL^{(0)}$). \\ [1mm]
%
Apart from plotting standard performance curves and reporting their AUC values per dataset in App. \ref{app:all_results}, we primarily rely on ranks to aggregate the performance of an AL method across datasets.
For each dataset and query size, the AUC values of all AL methods are sorted and assigned a rank based on position, with the best rank being 1.
These ranks can safely be averages across datasets as they are no longer subjected to scaling differences of each dataset.
Additionally, we employ Critical Difference (CD) diagrams (like Fig. \ref{fig:ranks_by_domain}) for statistical testing.
CD diagrams \cite{IsmailFawaz2018deep} use the Wilcoxon signed-rank test, which is a variant of the paired T-test, to find significant differences of ranks between AL methods.
For a detailed description of how every CD diagram is created, please refer to App. \ref{app:cd_diagrams}.

@@

Using additional resources, like excessive training time, or direct access to a labeled test set, an oracle method for AL finds the oracle set $\mathcal{O}_b$ for a given dataset, model, and

@@

However, due to the combinatorial nature of the problem, this is computationally

@@

annealing to search for a subset with maximal test accuracy and used the best solution

@@

uniformly samples at iteration $i$ a subset $\mathcal{U}_S$ of size $\tau$ of the not already

@@

point $u$ for which the classifier $\hat{y}^{(i)}$ trained on

@@

our oracle frequently encounters situations where every point in $u$ would

@@

defaults to margin sampling from the entire unlabeled pool $\mathcal{U}^{(i)}$ in that iteration.
The resulting greedy algorithm

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Implementation Details}\label{sec:implementation_details}
At each iteration $i$ the AL method picks an unlabeled datapoint based on a fixed set of information $\{\mathcal{L}^{(i)}, \mathcal{U}^{(i)}, B, |\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|, \text{acc}^{(i)}, \text{acc}^{(1)}, \hat y^{(i)}, \text{opt}_{\hat y}\}$, where $\text{opt}_{\hat y}$ is the optimizer used to fit $\hat y^{(i)}$.
This set grants full access to the labeled and unlabeled set, as well as all parameters of the classifier and the optimizer.
Additionally, we provide meta-information, like the size of the seed set through $|\mathcal{L}^{(i)}|-|\mathcal{L}^{(1)}|$, the remaining budget though the addition of $B$ and the classifiers potential through $\text{acc}^{(1)}$ and $\text{acc}^{(i)}$.
We allow AL methods to derive information from this set, e.g. predictions of the classifier $\hat y^{(i)}(x); \hspace{2mm} x \in \mathcal{U}^{(i)} \cup \mathcal{L}^{(i)}$, clustering, or even training additional models.
However, the method may not incorporate external information e.g. other datasets, queries to recover additional labels, additional training steps for $\hat y$, or the test/validation set. \\
For our study we selected AL methods with good performances reported by multiple different sources that can work with the set of information stated above.
For a list of all AL methods, please refer to Table \ref{tab:results}, with detailed descriptions being found in Appendix \ref{app:acquisition_functions}. \\ [1mm]
%
The model $\hat y^{(i)}$ can be trained in two ways. Either the parameters of the model are reset to a fixed initial setting $\hat y^{(0)}$ after each AL iteration and the classifier is trained from scratch with the updated labeled set $\mathcal{L}^{(i)}$, or the previous state $\hat y^{(i-1)}$ is retained and the classifier is fine-tuned on $\mathcal{L}^{(i)}$ for a reduced number of epochs.
In this work, we use the fine-tuning method for un-encoded datasets to save computational time, while we use the from-scratch training for encoded datasets since they have very small classifiers and this approach generally produces better results.
Our fine-tuning scheme always trains for at least one epoch and employs an aggressive early stopping with a patience of 2 afterwards.
%This work employs a validation split that is produced from the source dataset itself before active learning instead of splitting it from the seed set or omitting it altogether.
%Even though the use of a fully labeled validation set might be regarded as impractical, since such a set will never exist during deployment, we strongly advocate for using it in \textbf{research}, as it can be used for hyperparameter tuning, resulting in more stable hyperparameters and reducing the overall training stochasticity.
%We optimize all hyperparameters on our fully labeled validation sets using grid search.

@@

Please note, that we are averaging not only over runs, but also over query sizes per dataset, impacting AL methods that do not adapt well to a wide range of query sizes.

@@

As mentioned in contribution \ref{cont:domains}, our results on real-world data show significant differences in the performance of methods between data domains:

@@

Results for the semi-supervised domain appear mostly in line with the other 3 domains, but a closer analysis of performances split into encoded images and encoded tabular reveals the need for further research.

@@

Both types of methods have principled shortcomings in terms of their utilized information that makes them unsuitable for certain data distributions. 

@@

To avoid methods memorizing these datasets, they are generated from scratch for each experiment. \\

@@

The honeypot is located on the likely decision boundary of a classifier that is trained on the beneficial samples to maximize its adverse impact on purely uncertainty-based AL methods.
Diverging Sine samples datapoints for each class from two diverging sinusoidal functions that are originating from the same y-intercept.

@@

Both margin sampling and BADGE (the two best methods) being vulnerable to adverse samples or simply measurement noise, highlights the need for further research into robust AL methods. \\
